\documentclass[11pt, a4paper]{article}
\usepackage{../mysty}
\renewcommand{\lessontitle}{First-order logic, part~II}
\renewcommand{\fulltitle}{Homework 3}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% START DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}
\date{}



\begin{center}
{\Large {\bf \fulltitle}}
\end{center}
\hfill{\bf B04902012 Han Sheng, Liu}
\vspace{0.7cm}


\paragraph*{Question 1.}
Construct a structure $\cA$ and a formula $\varphi$ such that
$\cA \models \forall x \exists y \ \varphi$,
but $\cA \not\models \exists x \forall y \ \varphi$.
\begin{framed}
\solution

Consider $A=\{\T, \F\}$, $\varphi=x\approx y$. 

For every $x$, there must exist $y$ such that $x \approx y$.
Just simply pick $x$ for $y$.

However, there does not exist $x$ such that for every $y$, $x\approx y$.
If $y=\T$, then $x$ must be $\T$; if $y=\F$, then $x$ must be $\F$.
\end{framed}



\paragraph*{Question 2.}
There are many ways to ``encode''
propositional calculus in first-order logic.
One way is to view a propositional variable $p$ as
a relation symbol of arity $0$
as explained in the class.

Here is another way.
Let $L = \{c\}$ be a vocabulary that consists of only one
constant symbol $c$.
Given a propositional formula $\alpha$ 
using variables $p_1,\ldots,p_k$,
consider the FO formula $\Phi$
obtained from $\alpha$ by 
changing every $p_i$ with an atomic $x_i \approx c$.
For example, if $\alpha$ is 
$$p_1\wedge \neg p_2,$$
then $\Phi$ is 
$$
(x_1 \approx c) \ \wedge \ \neg (x_2 \approx c).
$$
Prove that
$\alpha$ is satisfiable (in the sense of propositional calculus)
if and only if $\Phi$ is.

\begin{framed}
\solution

    Note that $\Phi$ and $\alpha$ are actually the same after replacing those $(x_i\approx c)$ with $p_i$ in $\Phi$.

    Also note that satisfying assignments and models have the same recursive definition.

\noindent$\Rightarrow$:

    $\alpha$ is satisfiable.
    Thus, there exists an assignment $w$ that satisfies $\alpha$.

    Now consider the following interpretation $(\cA,\val)$.

    $A=\{\T, \F\}$, $c=\T$, $\val(x_i)=w(p_i)$.

    Then $(\cA,\val)\models(x_i\approx c)$ if and only if $w(p_i)=\T$.

    Thus, $w\models\alpha\Rightarrow(\cA, \val)\models\Phi$, because satisfying assignments and models have the same recursive definition.

\noindent$\Leftarrow$:

    $\Phi$ is satisfiable.
    Thus, there exists an interpretation $(\cA, \val)$ satisfying $\Phi$.

    Now consider the following assignment $w$ for $\alpha$.

    $w(p_i)=\T$ if and only if $(\cA, \val)\models (x_i\approx c)$.

    Thus, $(\cA, \val)\models \Phi \Rightarrow w \models \alpha$, because satisfying assignments and models have the same recursive definition.
\end{framed}

\paragraph*{Question 3.}
Consider a sentence of the following form:
\begin{eqnarray*}
\Phi & := & \exists x_1 \cdots \exists x_n \ \forall y_1 \cdots \forall y_m \ \varphi,
\end{eqnarray*}
where $\varphi$ is quantifier free and does {\em not} contain any function and constant symbol, and $n,m \geq 1$.

Prove that if $\Phi$ is satisfiable, then there is a structure $\cA$ that satisfies $\Phi$ with $|A|\leq n$. 
\begin{framed}
\solution

$\Phi$ is a sentence, so each variable in $\Phi$ is one of the $x_i$ or $y_i$.

Let $\Phi'= \forall y_1 \cdots \forall y_m \ \varphi$, then every model of $\Phi$ is a model of $\Phi'$, and vice versa.

Assume that $\Phi'$ is satisfiable, then there is a interpretation $(\cA', \val)\models\Phi'$.

Let $X=\{\val(x_i)\}$. Consider the substructure $\cA$ such that $A=A'\cap X$. It is obvious that $|A|\leq n$. Now prove that $(\cA, \val)\models \Phi'$

\begin{align*}
(\cA', \val) \models \Phi' &\Rightarrow \forall\ Y\in A'^{m}, (\cA', \val[\bar{y}\mapsto Y])\models \phi\\
                           &\Rightarrow \forall\ Y\in A^{m}, (\cA, \val[\bar{y}\mapsto Y])\models \phi \text{ (because $A\subseteq A'$)}\\
                           &\Rightarrow (\cA, \val) \models \Phi'\\
                           &\Rightarrow (\cA, \val) \models \Phi
\end{align*}
\end{framed}


\paragraph*{Question 4.}
Let $h:\cA \to \cB$ be a homomorphism.
Define a relation $\sim_h$ on $A$ as follows:
$a \sim_h a'$ if and only if $h(a)=h(a')$.
Prove that $\sim_h$ is an equivalence relation on $A$,
and that $\sim_h$ is, in fact, a congruence in $\cA$.

\begin{framed}
\solution

\begin{itemize}
\item Reflexive

$\forall a\in A$, $h(a)=h(a)$.

Thus, $\forall a\in A$, $a\sim_ha$.

\item Symmetric
\begin{flalign*}
\forall a,\ b\in A,\ a\sim_hb&\Rightarrow h(a)=h(b)\\
        &\Rightarrow h(b)=h(a)\\
        &\Rightarrow b\sim_ha
\end{flalign*}

\item Transitive
\begin{flalign*}
\forall a,\ b,\ c\in A,\ a\sim_hb,\ b\sim_hc&\Rightarrow h(a)=h(b),\ h(b)=h(c)\\
                   &\Rightarrow h(a)=h(c)\\
                   &\Rightarrow a\sim_hc
\end{flalign*}
\item Congruence
\begin{flalign*}
\forall a,\ b\in A,\ a\sim_hb&\Rightarrow h(a)=h(b)\\
                             &\Rightarrow f(h(a))=f(h(b))\\
                             &\Rightarrow h(f(a))=h(f(b))\\
                             &\Rightarrow f(a)\sim_hf(b)
\end{flalign*}
\end{itemize}
\end{framed}

\paragraph*{Question 5.}
Let $h:\cA\to\cB$ be a strong and surjective homomorphism,
and let $\cA/\!\!\sim_h$ be the factor of $\cA$ modulo $\sim_h$.
Define a function $\xi:A/\!\!\sim_h \to B$,
where $\xi([a]_{\sim_h}) = h(a)$.
Prove that $\xi$ is an isomorphism.

\begin{framed}
\solution
Let $\cC=\cA/\sim_h$.
\begin{itemize}
\item $\xi$ is a homomorphism

$\forall R,\ f,\ c\in L$,

\begin{flalign*}
\forall \bar{a}\in A^{ar(f)}\, \xi(f^\cC([\bar{a}]_{\sim_h})&=\xi([f^\cA(\bar{a})]_{\sim_h})\\
                                                            &=h(f^\cA(\bar{a}))\\
                                                            &=f^\cB(h(\bar{a}))\\
                                                            &=f^\cB(\xi([\bar{a}]_{\sim_h}))

\end{flalign*}
\end{itemize}



\end{framed}



\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% END OF DOCUMENT %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

